Revenue Maximization for Buyers with Costly Participation

### 2025-03-19 paper/paper.tex
Apply cleverref

\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{thmtools}
\usepackage{hyperref}
\usepackage{pgf, tikz}
\usepackage{xspace, units}
\usepackage{typearea}
\allowdisplaybreaks

@@

@@

 In this article based on the work of \citet{primary}, we derive a model in \autoref{sec:model} that incorporates private participation costs as well as valuations.
 Next we focus on finding a formulation of the expected revenue that is to be maximized as also done by \citet{myerson} in \autoref{sec:incentive-comptaible-mechanism-revenue}.
 We will see that this problem becomes inherently non-convex and thus we focus on finding an FPTAS that achieves a $\mathrm{OPT} - 4\epsilon$ approximation where $\mathrm{OPT}$ refers to the optimal single buyer mechanism in \autoref{sec:revenue-maximization}.
\end{section}

@@

 For the first part, in common single parameter mechanisms \citet{myerson} found a clear characterization of a truthful mechanism (\autoref{lemma:myersons-lemma}).

@@

 \citet{primary} found that we can base our work on Myerson's Lemma (\ref{lemma:myersons-lemma})
 by generalizing a truthful single-parameter mechanism to a one that takes participation cost into account (\autoref{lemma:thruthful-mechanism}).

@@

 Next to maximize the expected revenue, it is convenient to have a formulation for it.
 In the single parameter setting, again \citet{myerson} provided such as well (\autoref{lemma:myerson-revenue}).

@@

 Following from that we derive an expression for the expected revenue given private participation costs.
 For that we first formulize the mechanism described in \autoref{lemma:thruthful-mechanism}.

@@

 \begin{align*}
     v_x(c) = \inf_{v \in \RR_{\geq 0}} \{\underbrace{vx(v) - p(v) \geq c}_{u^\mathcal{M}(v)}\}
 \end{align*}
 Next we can define the allocation function $x_c$ as simply implementing the characterization of \autoref{lemma:thruthful-mechanism},
 by allocating $x(v)$ if the buyer has a value that is at least $v_x(c)$ and $0$ otherwise.

 \begin{align*}
     x_c(v) = \begin{cases}
                  x(v) & v \geq v_x(c)    \\
                  0    & \text{otherwise}
              \end{cases}
 \end{align*}
 Similar the payment rule $p_c(v)$ delegates the payment rule $p$ in the case that $v \geq v_x(c)$ and otherwise expects no payment.
 Together, they define the mechanism $\mathcal{M}_c = (x_c, p_c)$ induced by the mechanism $\mathcal{M}$.
 See \autoref{fig:visualization-of-induced-mechanism} for a visualization of the allocation rule.

@@

 If we now define a virtual value function that is conditional on $c$, by using the conditional cumulative distribution $\bar{F}_c$ and the conditional density function $\hat{f}_c$ as
 $\varphi_c(v) = v - \frac{1- \bar{F}_c(v)}{\hat{f}_c(v)}$ then we can express the expected revenue in \autoref{theorem:expected-revenue}.

@@

     Consider the case that $c > -p_0$. Note that $-p_0$ corresponds to a payment the buyer would get from the mechanism for participating.
     That means in the case that $c > -p_0$, we do not simply always participate, since our utility could be negative. Therefore, we also have $v_x(c) > 0$.
     We first consider the case that we participate, i.e. $v \geq v_x(c)$.
     In that case the buyer has to pay $p(v)$.
     \begin{align}
         \label{eq:payment-participating}
         p_c(v) = p(v) = v x(v) - \int_0^v x(t) dt + p_0
     \end{align}
     We use linearity of the integral and the definition of utility to express this in terms of the allocation function $x_c$.
     \begin{equation}
         \label{eq:utility-participating}
         \begin{aligned}
             \int_0^v x(t) dt & = \int_0^{v_c(x)} x(t) dt + \int_{v_c(x)}^v x(t) dt     \\
                              & = u^\mathcal{M}(v_c(x)) + p_0 + \int_{v_c(x)}^v x(t) dt \\
                              & = c + p_0 + \int_{v_c(x)}^v x(t) dt
         \end{aligned}
     \end{equation}
     Here the last step uses that $u^\mathcal{M}(v_c(x)) = c$ per definition of $v_c(x)$.
     Plugging \autoref{eq:utility-participating} into \autoref{eq:payment-participating} yields \autoref{eq:payment-participating-final}:
     \begin{align}
         \label{eq:payment-participating-final}
         p_c(v) = v x_c(v) - \int_0^v x_c(t) dt - c
     \end{align}
     Next we consider the case that the buyer does not participate, i.e. $v < v_x(c)$.
     In that case, the payment rule requests no payment, i.e. $p_c(v) = 0$, yet conveniently some part of the expression found before are also $0$:
     \begin{align*}
         p_c(v) = 0 = v x_c(v) - \int_0^v x_c(t) dt
     \end{align*}
     Thus, we can represent this entire case using an indicator function:
     \begin{align*}
         p_c(v) = 0 = v x_c(v) - \int_0^v x_c(t) dt - c \1{v \geq v_x(c)}
     \end{align*}
     Drawing the expectation around the value $v \sim \bar{F}_c$ while keeping the participation cost still fixed yields:
     \begin{equation}
         \label{eq:expected-revenue-first-case}
         \begin{aligned}
             \Ex[v \sim \bar{F}_c]{p_c(v)} & = \Ex[v \sim \bar{F}_c]{v x_c(v) - \int_0^v x_c(t) dt - c \1{v \geq v_x(c)}}   \\
                                           & = \Ex[v \sim \bar{F}_c]{v x_c(v) - \int_0^v x_c(t) dt} - \Pr{v \geq v_x(c)}c   \\
                                           & = \Ex[v \sim \bar{F}_c]{x_c(v) \varphi_c(v)} - (1 - \bar{F}_c(v_x(c))) \cdot c \\
         \end{aligned}
     \end{equation}
     Here in the last step we used that for fixed $c$, $\mathcal{M}_c$ is a truthful mechanism thus we can apply \autoref{lemma:myerson-revenue}, simultaneously we use the definition of the cumulative distribution.

@@

     Now we can finally draw the expectation around the previously fixed $c$, combine \autoref{eq:expected-revenue-first-case} and \autoref{eq:expected-revenue-second-case} to get the wished theorem:

@@

 This leads us to a problem definition for which an optimal mechanism is computable in polynomial time with respect to $\frac{1}{\epsilon}$.
 We will see later how to do this.
 Still it is not clear that this mechanism is close to the optimal mechanism for the original problem, here \autoref{theorem:approximation-guarantee} comes into play.

@@

 \begin{proof}[Proof of \autoref{theorem:approximation-guarantee}.]
     For any $\bar{F}$, let $\mathcal{M}'$ be the optimal mechanism on the discretized type distribution $\bar{F}'$ for which the allocation rule is also discretized.
     First, since we defined $\bar{F}'$ as the \textit{rounded down} type distribution, for any mechanism $\mathcal{M}$ with monotone allocation rule holds, that $\mathrm{Rev}(\bar{F}, \mathcal{M}) \geq \mathrm{Rev}(\bar{F}', \mathcal{M})$, so in particular this also applies to $\mathcal{M}'$.

     Next, we define $\widehat{\mathcal{M}}$ as the optimal mechanism on $\bar{F}'$ (one that is allowed to allocate continuous values).
     We use \autoref{lemma:allocation-function-within-e}, since our allocation function $x'$ of mechanism $\mathcal{M}'$
     allocates within an $\epsilon$-interval of the allocation function $\hat{x}$ of $\widehat{\mathcal{M}}$.
     Thus, according to \autoref{lemma:allocation-function-within-e}, we got $\mathrm{Rev}(\bar{F}', \mathcal{M}') \geq \mathrm{Rev}(\bar{F}', \widehat{\mathcal{M}}) - \epsilon$.

     Finally, note, that for the two type distributions $\bar{F}$ and $\bar{F}'$ it holds that they originate from the same probability measure, yet the value of the random variables differs by at most $\epsilon$.
     Also note, that we did not round the participation cost, thus the difference with respect to the infinity-norm $|| \cdot ||_{\infty}$ is also at most $\epsilon$.
     Therefore, we can apply \autoref{lemma:difference-in-optimal-mechanisms}, since $\widehat{\mathcal{M}}$ is optimal mechanism on $\bar{F}'$,
     to get $\mathrm{Rev}(\bar{F}', \widehat{\mathcal{M}}) \geq \mathrm{OPT}(\bar{F}') - 3\epsilon$.

@@

 We will use a dynamic program based on the approximation degree $\epsilon$ to calculate for all possible valuations
 an allocation function that optimizes the revenue.
 We will make sure this allocation rule is non-decreasing and will follow the payment rule through \autoref{lemma:thruthful-mechanism}
 in order to also get a truthful mechanism.
 Note that \autoref{lemma:thruthful-mechanism} also tells us, that it is enough to find a single parameter allocation rule and optimize the revenue of the induced mechanism $\mathcal{M}_c$ with payment rule $p_c$.

@@

 See \autoref{fig:fptas-recursion-formular} for a visualization of this theorem.

@@

         \node at (0, 0) [below left] {0};
         \foreach \x in {1, 2, 3, 3, 4} {
                 \node at (0, \x) [left] {\x};
                 \node at (\x, 0) [below] {\x};
             }
     \end{tikzpicture}
     \caption{
         Visualization of \autoref{theorem:recursion-formular}. For $R(3,2,k)$ we need to maximize over all $R(2, j', k-j)$ where $j' \leq j = 2$.
     }
     \label{fig:fptas-recursion-formular}
 \end{figure}

@@

     Using dynamic programming by applying \autoref{theorem:recursion-formular} and fixing the base cases $R(1,j,k) = 0$ for any $k \leq j$ and $R(1,j,k)=-\infty$ for any $k > j$,
     we can calculate for all $\frac{1}{\epsilon^4}$ combinations of $i,j$ and $k$ the respective values of $R(i,j,k)$.
     Each of those $R(i,j,k)$ needs at most $\frac{1}{\epsilon}$ calculations because of the maximum.
     Therefore, we calculate the revenue of $\mathcal{M}'$ in $\mathcal{O}\left(\frac{1}{\epsilon^5}\right)$ time.
     Recovering the actual mechanism $\mathcal{M}'$ and its allocation rule from that is simply a matter of backtracking the $j'$ that maximized our expression.
 \end{proof}

 All that is left is to proof \autoref{theorem:recursion-formular}.
 \begin{proof}[Proof of \autoref{theorem:recursion-formular}]
     First recall the definition of $R(i,j,k)$ (\autoref{eq:rijk}):
     \begin{align*}
         R(i,j,k) = \max_{\mathcal{M}: x(i\epsilon) = j\epsilon, u(i\epsilon) = k \epsilon^2} \Pr{v \leq i\epsilon}\Ex{p_c(v)\mid v \leq i \epsilon}
     \end{align*}
     Observe that we can use the linearity of expectation to split up the expectation of \autoref{eq:rijk} into the case that $v \leq (i-1)\epsilon$ and the case that $v = i \epsilon$.

@@

     Note that this is similar to \autoref{eq:rijk} with $v = (i-1) \epsilon$, only the maximization constraints differ.
     For that we only need to fix $x((i-1)\epsilon)$ and $u((i-1)\epsilon)$.
     Conveniently, we can express the latter in terms of things that are already fixed.
     Note that we can write $u(i\epsilon)$ as follows:

@@

 \subsection{Reduction to Multiple Buyer}
 Using the mechanism designed in \autoref{sec:revenue-maximization} for a single buyer, \citet{primary} derive an approximate algorithm for multiple buyers.
 For that they use that any algorithm allocating a single item to multiple buyers allocates the item in expectation to at most once.
 Therefore, the sum of ex-ante probabilities for allocating it to each buyer is also at most $1$.
 One now splits the probabilities into shares for each buyer while maximizing the revenue.
 After that each buyer is approached sequentially with the single parameter mechanism from \autoref{sec:revenue-maximization}.
 The mechanism is modified, s.t. the item is only allocated with the precomputed probability to each buyer.
 This general reduction scheme was first proposed by \citet{alaei2012bayesian,alaei2014bayesian} leads to a $\frac{1}{2}\mathrm{OPT} - \epsilon$ approximation of the optimal mechanism in $\mathrm{poly}\left(\frac{1}{\epsilon}\right)$ computation time in our case.
\end{section}

>>>>

\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{thmtools}
\usepackage{pgf, tikz}
\usepackage{xspace, units}
\usepackage{typearea}
\allowdisplaybreaks

@@

\usepackage{hyperref}
\usepackage[noabbrev]{cleveref}

@@

 In this article based on the work of \citet{primary}, we derive a model in \cref{sec:model} that incorporates private participation costs as well as valuations.
 Next we focus on finding a formulation of the expected revenue that is to be maximized as also done by \citet{myerson} in \cref{sec:incentive-comptaible-mechanism-revenue}.
 We will see that this problem becomes inherently non-convex and thus we focus on finding an FPTAS that achieves a $\mathrm{OPT} - 4\epsilon$ approximation where $\mathrm{OPT}$ refers to the optimal single buyer mechanism in \cref{sec:revenue-maximization}.
\end{section}

@@

 For the first part, in common single parameter mechanisms \citet{myerson} found a clear characterization of a truthful mechanism (\Cref{lemma:myersons-lemma}).

@@

 \citet{primary} found that we can base our work on Myerson's lemma (\ref{lemma:myersons-lemma})
 by generalizing a truthful single-parameter mechanism to a one that takes participation cost into account (\Cref{lemma:thruthful-mechanism}).

@@

 Next to maximize the expected revenue, it is convenient to have a formulation for it.
 In the single parameter setting, again \citet{myerson} provided such as well (\Cref{lemma:myerson-revenue}).

@@

 Following from that we derive an expression for the expected revenue given private participation costs.
 For that we first formulize the mechanism described in \Cref{lemma:thruthful-mechanism}.

@@

 \begin{align*}
     v_x(c) = \inf_{v \in \RR_{\geq 0}} \{\underbrace{vx(v) - p(v) \geq c}_{u^\mathcal{M}(v)}\}
 \end{align*}
 Next we can define the allocation function $x_c$ as simply implementing the characterization of \Cref{lemma:thruthful-mechanism},
 by allocating $x(v)$ if the buyer has a value that is at least $v_x(c)$ and $0$ otherwise.

 \begin{align*}
     x_c(v) = \begin{cases}
                  x(v) & v \geq v_x(c)    \\
                  0    & \text{otherwise}
              \end{cases}
 \end{align*}
 Similar the payment rule $p_c(v)$ delegates the payment rule $p$ in the case that $v \geq v_x(c)$ and otherwise expects no payment.
 Together, they define the mechanism $\mathcal{M}_c = (x_c, p_c)$ induced by the mechanism $\mathcal{M}$.
 See \cref{fig:visualization-of-induced-mechanism} for a visualization of the allocation rule.

@@

 If we now define a virtual value function that is conditional on $c$, by using the conditional cumulative distribution $\bar{F}_c$ and the conditional density function $\hat{f}_c$ as
 $\varphi_c(v) = v - \frac{1- \bar{F}_c(v)}{\hat{f}_c(v)}$ then we can express the expected revenue in \cref{theorem:expected-revenue}.

@@

     Consider the case that $c > -p_0$. Note that $-p_0$ corresponds to a payment the buyer would get from the mechanism for participating.
     That means in the case that $c > -p_0$, we do not simply always participate, since our utility could be negative. Therefore, we also have $v_x(c) > 0$.
     We first consider the case that we participate, i.e. $v \geq v_x(c)$.
     In that case the buyer has to pay $p(v)$.
     \begin{align}
         \label{eq:payment-participating}
         p_c(v) = p(v) = v x(v) - \int_0^v x(t) dt + p_0
     \end{align}
     We use linearity of the integral and the definition of utility to express this in terms of the allocation function $x_c$.
     \begin{equation}
         \label{eq:utility-participating}
         \begin{aligned}
             \int_0^v x(t) dt & = \int_0^{v_c(x)} x(t) dt + \int_{v_c(x)}^v x(t) dt     \\
                              & = u^\mathcal{M}(v_c(x)) + p_0 + \int_{v_c(x)}^v x(t) dt \\
                              & = c + p_0 + \int_{v_c(x)}^v x(t) dt
         \end{aligned}
     \end{equation}
     Here the last step uses that $u^\mathcal{M}(v_c(x)) = c$ per definition of $v_c(x)$.
     Plugging \cref{eq:utility-participating} into \cref{eq:payment-participating} yields \cref{eq:payment-participating-final}:
     \begin{align}
         \label{eq:payment-participating-final}
         p_c(v) = v x_c(v) - \int_0^v x_c(t) dt - c
     \end{align}
     Next we consider the case that the buyer does not participate, i.e. $v < v_x(c)$.
     In that case, the payment rule requests no payment, i.e. $p_c(v) = 0$, yet conveniently some part of the expression found before are also $0$:
     \begin{align*}
         p_c(v) = 0 = v x_c(v) - \int_0^v x_c(t) dt
     \end{align*}
     Thus, we can represent this entire case using an indicator function:
     \begin{align*}
         p_c(v) = 0 = v x_c(v) - \int_0^v x_c(t) dt - c \1{v \geq v_x(c)}
     \end{align*}
     Drawing the expectation around the value $v \sim \bar{F}_c$ while keeping the participation cost still fixed yields:
     \begin{equation}
         \label{eq:expected-revenue-first-case}
         \begin{aligned}
             \Ex[v \sim \bar{F}_c]{p_c(v)} & = \Ex[v \sim \bar{F}_c]{v x_c(v) - \int_0^v x_c(t) dt - c \1{v \geq v_x(c)}}   \\
                                           & = \Ex[v \sim \bar{F}_c]{v x_c(v) - \int_0^v x_c(t) dt} - \Pr{v \geq v_x(c)}c   \\
                                           & = \Ex[v \sim \bar{F}_c]{x_c(v) \varphi_c(v)} - (1 - \bar{F}_c(v_x(c))) \cdot c \\
         \end{aligned}
     \end{equation}
     Here in the last step we used that for fixed $c$, $\mathcal{M}_c$ is a truthful mechanism thus we can apply \cref{lemma:myerson-revenue}, simultaneously we use the definition of the cumulative distribution.

@@

     Now we can finally draw the expectation around the previously fixed $c$, combine \cref{eq:expected-revenue-first-case} and \cref{eq:expected-revenue-second-case} to get the wished theorem:

@@

 This leads us to a problem definition for which an optimal mechanism is computable in polynomial time with respect to $\frac{1}{\epsilon}$.
 We will see later how to do this.
 Still it is not clear that this mechanism is close to the optimal mechanism for the original problem, here \cref{theorem:approximation-guarantee} comes into play.

@@

 \begin{proof}[Proof of \Cref{theorem:approximation-guarantee}.]
     For any $\bar{F}$, let $\mathcal{M}'$ be the optimal mechanism on the discretized type distribution $\bar{F}'$ for which the allocation rule is also discretized.
     First, since we defined $\bar{F}'$ as the \textit{rounded down} type distribution, for any mechanism $\mathcal{M}$ with monotone allocation rule holds, that $\mathrm{Rev}(\bar{F}, \mathcal{M}) \geq \mathrm{Rev}(\bar{F}', \mathcal{M})$, so in particular this also applies to $\mathcal{M}'$.

     Next, we define $\widehat{\mathcal{M}}$ as the optimal mechanism on $\bar{F}'$ (one that is allowed to allocate continuous values).
     We use \cref{lemma:allocation-function-within-e}, since our allocation function $x'$ of mechanism $\mathcal{M}'$
     allocates within an $\epsilon$-interval of the allocation function $\hat{x}$ of $\widehat{\mathcal{M}}$.
     Thus, according to \cref{lemma:allocation-function-within-e}, we got $\mathrm{Rev}(\bar{F}', \mathcal{M}') \geq \mathrm{Rev}(\bar{F}', \widehat{\mathcal{M}}) - \epsilon$.

     Finally, note, that for the two type distributions $\bar{F}$ and $\bar{F}'$ it holds that they originate from the same probability measure, yet the value of the random variables differs by at most $\epsilon$.
     Also note, that we did not round the participation cost, thus the difference with respect to the infinity-norm $|| \cdot ||_{\infty}$ is also at most $\epsilon$.
     Therefore, we can apply \cref{lemma:difference-in-optimal-mechanisms}, since $\widehat{\mathcal{M}}$ is optimal mechanism on $\bar{F}'$,
     to get $\mathrm{Rev}(\bar{F}', \widehat{\mathcal{M}}) \geq \mathrm{OPT}(\bar{F}') - 3\epsilon$.

@@

 We will use a dynamic program based on the approximation degree $\epsilon$ to calculate for all possible valuations
 an allocation function that optimizes the revenue.
 We will make sure this allocation rule is non-decreasing and will follow the payment rule through \cref{lemma:thruthful-mechanism}
 in order to also get a truthful mechanism.
 Note that \cref{lemma:thruthful-mechanism} also tells us, that it is enough to find a single parameter allocation rule and optimize the revenue of the induced mechanism $\mathcal{M}_c$ with payment rule $p_c$.

@@

 See \cref{fig:fptas-recursion-formular} for a visualization of this theorem.

@@

         \node at (0, 0) [below left] {0};
         \foreach \x in {1, 2, 3, 3, 4} {
                 \node at (0, \x) [left] {\x};
                 \node at (\x, 0) [below] {\x};
             }
     \end{tikzpicture}
     \caption{
         Visualization of \Cref{theorem:recursion-formular}. For $R(3,2,k)$ we need to maximize over all $R(2, j', k-j)$ where $j' \leq j = 2$.
     }
     \label{fig:fptas-recursion-formular}
 \end{figure}

@@

     Using dynamic programming by applying \cref{theorem:recursion-formular} and fixing the base cases $R(1,j,k) = 0$ for any $k \leq j$ and $R(1,j,k)=-\infty$ for any $k > j$,
     we can calculate for all $\frac{1}{\epsilon^4}$ combinations of $i,j$ and $k$ the respective values of $R(i,j,k)$.
     Each of those $R(i,j,k)$ needs at most $\frac{1}{\epsilon}$ calculations because of the maximum.
     Therefore, we calculate the revenue of $\mathcal{M}'$ in $\mathcal{O}\left(\frac{1}{\epsilon^5}\right)$ time.
     Recovering the actual mechanism $\mathcal{M}'$ and its allocation rule from that is simply a matter of backtracking the $j'$ that maximized our expression.
 \end{proof}

 All that is left is to proof \cref{theorem:recursion-formular}.
 \begin{proof}[Proof of \Cref{theorem:recursion-formular}]
     First recall the definition of $R(i,j,k)$ (\cref{eq:rijk}):
     \begin{align*}
         R(i,j,k) = \max_{\mathcal{M}: x(i\epsilon) = j\epsilon, u(i\epsilon) = k \epsilon^2} \Pr{v \leq i\epsilon}\Ex{p_c(v)\mid v \leq i \epsilon}
     \end{align*}
     Observe that we can use the linearity of expectation to split up the expectation of \cref{eq:rijk} into the case that $v \leq (i-1)\epsilon$ and the case that $v = i \epsilon$.

@@

     Note that this is similar to \cref{eq:rijk} with $v = (i-1) \epsilon$, only the maximization constraints differ.
     For that we only need to fix $x((i-1)\epsilon)$ and $u((i-1)\epsilon)$.
     Conveniently, we can express the latter in terms of things that are already fixed.
     Note that we can write $u(i\epsilon)$ as follows:

@@

 \subsection{Reduction to Multiple Buyer}
 Using the mechanism designed in \cref{sec:revenue-maximization} for a single buyer, \citet{primary} derive an approximate algorithm for multiple buyers.
 For that they use that any algorithm allocating a single item to multiple buyers allocates the item in expectation to at most once.
 Therefore, the sum of ex-ante probabilities for allocating it to each buyer is also at most $1$.
 One now splits the probabilities into shares for each buyer while maximizing the revenue.
 After that each buyer is approached sequentially with the single parameter mechanism from \cref{sec:revenue-maximization}.
 The mechanism is modified, s.t. the item is only allocated with the precomputed probability to each buyer.
 This general reduction scheme was first proposed by \citet{alaei2012bayesian,alaei2014bayesian} leads to a $\frac{1}{2}\mathrm{OPT} - \epsilon$ approximation of the optimal mechanism in $\mathrm{poly}\left(\frac{1}{\epsilon}\right)$ computation time in our case.
\end{section}
